---
title: "covid 19"
author: "Nathan Abromovich"
date: "Fall 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

`
```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
  

This plot is showing that a very small percentage of people have died from COVID.  Of the 1831 females that got sick only 25 died.  In men that caught COVID, a total of 1141, 44 have died of the virus. With this data we can see that males die more often than females. 



  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The chi sq test provide the exact numbers. It shoes that 1831 people were females and that 1411 males that are alive.  It also shows that 25 females and 44 males have died.  The Fisher test shows us the odds of dying for male is that they are 2.3 times more likely to die than females. 


## Conclusions
In conclusion, the data shows men were more likely to die than females for the COVID virus. The data used in this study was from the beginning the start of COVID, March 2020.  Therefor these results could continue to change. 
